---
title: "Predicting Genetic Disorders"
author:
  Emma Oo^[University of San Diego, eoo@sandiego.edu], Sindhu Bhattarai^[University of San Diego, sbhattarai@sandiego.edu], Dave Friesen^[University of San Diego, dfriesen@sandiego.edu]
date: "06/27/2022"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: rmarkdown::html_vignette
---

```{r setup, echo = FALSE, message = FALSE}
# Load R libraries
library(caret)
library(DescTools)
library(e1071)

# Expand output width and minimize exp notation
options(width = 150)
options(scipen = 100)
options(digits = 1)

# Set style defaults
knitr::opts_chunk$set(class.source = "source")
knitr::opts_chunk$set(class.output = "output")
knitr::opts_chunk$set(fig.width = 6.5, fig.height = (6.5 * .7), fig.align = "center")
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NA)
```

### Data Load and Validation

```{r }
gd_df = read.csv(file="/Users/sakshyamdahal/Desktop/MS_Data_Science/Applied Predictive Modeling/Final_project/train_genetic_disorders.csv", header= TRUE)
dim(gd_df)
```

###### Data Structure Review

```{r data_structure}
# Summarize base dataset and [optionally] sample rows
str(gd_df)
#head(gd_df, 3)
```




###### Initial feature reduction

Uninformative Feature Reduced

```{r prelim_feature_reduction}
# Define n/a columns and subset dataframe; Note retaining "some" informational variables like "Institute.Name" for
#   possible descriptive analytic purposes
drop_cols <- c("Patient.Id",
               "Patient.First.Name",
               "Family.Name",
               "Father.s.name",
               "Institute.Name",
               "Location.of.Institute",
               "Status",
               "Test.1",
               "Test.2",
               "Test.3",
               "Test.4",
               "Test.5",
               "Parental.consent",
               "Place.of.birth")
gd_df <- gd_df[ , !(names(gd_df) %in% drop_cols)]

dim(gd_df)
```

###### Class Target and Label Review

```{r class_target_label_review}
# Check for missing labels; set aside where missing
missing_target <- which(is.na(gd_df$Disorder.Subclass) | (gd_df$Disorder.Subclass == ""))
cat("Rows pre-subset for missing labels: ", format(nrow(gd_df), format = "d", big.mark = ","), sep = "")

gd_hold_df <- gd_df[missing_target, ]
gd_df <- gd_df[-missing_target, ]

cat("Deleted rows with missing labels: ", format(nrow(gd_hold_df), format = "d", big.mark = ","), sep = "")

cat(" Remaining rows (labeled): ", format(nrow(gd_df), format = "d", big.mark = ","), sep = "")
```


```{r frequency_target_class_review}
# Show frequency distribution for [prospective] target class(es)
show_frequency <- function(desc, c) {
  t <- as.data.frame(prop.table(table(c)))
  colnames(t) <- c("Class", "Frequency")
  cat(desc, "\n"); print(t[order(-t$Freq, t$Class), 1:2], row.names = FALSE)
}
show_frequency("Pre-Split Frequency Distribution", gd_df$Disorder.Subclass)

# Move the target class to "top" of dataframe so column removals don't impact
gd_df <- gd_df[ , c(ncol(gd_df), 1:(ncol(gd_df) - 1))]
target_col = 1


gd_df$Disorder.Subclass <- gsub("'", "", gd_df$Disorder.Subclass, fixed = TRUE)
gd_df$Disorder.Subclass <- gsub(" ", ".", gd_df$Disorder.Subclass, fixed = TRUE)
gd_df$Disorder.Subclass <- gsub("-", ".", gd_df$Disorder.Subclass, fixed = TRUE)

```


### Data Partition 

###### Split the dataframe as per target class (Disorder.Subclass vector)

**There is certain class imbalance problem. We will split the data frame using createDataPartition function using target class(Disorder.Subclass) which helps us to get even balance of all classes in both train and test data**

```{r data_splitting}
# Split data 80/20 train/test, using caret's inherent stratified split to compensate for class imbalance
set.seed(1)
train_index <- createDataPartition(gd_df$Disorder.Subclass, times = 1, p = 0.80, list = FALSE)
train_df <- gd_df[train_index, ]
test_df <- gd_df[-train_index, ]
show_frequency("Post-Split Frequency Distribution (Train)", train_df$Disorder.Subclass)
```

### Handling Missing Values

###### Checking  Mislabeled Data and  Missing Values

```{r mislabeled_data}
library(dplyr)
library(questionr)

#lapply(train_df, unique)

# changing values which are labeled as not available and no record to a readable format 'NA'
train_df[ train_df == "Not available"] <- NA
test_df[ test_df == "Not available"] <- NA

train_df[train_df == "No record"] <- NA
test_df[test_df == "No record"] <- NA

#changing None to "No" (there is None in one vector "Autopsy.shows.birth.defect..if.applicable." where none birth defects resemble no class)
train_df[train_df == "None"] <- "No"
test_df[test_df == "None"] <- "No"


freq.na(train_df)
sum(is.na(train_df))
```

**Note:We will be removing  "Birth.asphyxia" vector from our data as there is 46% missing data. Here, the imputation might create bias model.Further,we will be performing median imputation for integer vectors. For categorical vectors we will be encoding the missing value as "not provided". Here we can also do mode imputation but we would like to learn if the missing value has any relation with target. For numeric vectors we will be applying mean imputation.** 

```{r missing_values}
train_df <- subset(train_df, select = -c(Birth.asphyxia))
test_df <- subset(test_df, select = -c(Birth.asphyxia))

# Impute basic integer values with medians

medianf <- function(x) {
 result <- median(x, na.rm = TRUE)
 if (is.integer(x))
   result <- as.integer(result)
 return(result)
}
median_cols = c("Patient.Age", "Mother.s.age", "Father.s.age", "No..of.previous.abortion")
for (n in median_cols) {
  train_df[n][is.na(train_df[n])] <- apply(train_df[n], 2, medianf)
  test_df[n][is.na(test_df[n])] <- apply(test_df[n], 2, medianf)
}
                                           
# Impute categorical blanks with common "notprovided"; note we could also impute these with categorical mode,
#   or most frequent categorical value of each column using the cmode() function below
cols_tofill <- c("Inherited.from.father",
                 "Maternal.gene",
                 "Respiratory.Rate..breaths.min.",
                 "Heart.Rate..rates.min",
                 "Follow.up",
                 "Gender",
                 "Autopsy.shows.birth.defect..if.applicable.",
                 "Folic.acid.details..peri.conceptional.",
                 "H.O.serious.maternal.illness",
                 "H.O.radiation.exposure..x.ray.",
                 "H.O.substance.abuse",
                 "Assisted.conception.IVF.ART",
                 "History.of.anomalies.in.previous.pregnancies",
                 "Birth.defects",
                 "Blood.test.result",
                 "Genetic.Disorder")
train_df[cols_tofill][train_df[cols_tofill] == ""] <- "notprovided"
test_df[cols_tofill][test_df[cols_tofill] == ""] <- "notprovided"

train_df[cols_tofill][train_df[cols_tofill] == "-"] <- "notprovided"
test_df[cols_tofill][test_df[cols_tofill] == "-"] <- "notprovided"


cmode <- function(x) {
  uniqx <- unique(na.omit(x))
  uniqx[which.max(tabulate(match(x, uniqx)))]
}

# Impute what appear to be masked "flag" columns iwth placeholder -1 values. . .
flag_cols <- c("Symptom.1", "Symptom.2", "Symptom.3", "Symptom.4", "Symptom.5")
train_df[flag_cols][is.na(train_df[flag_cols])] <- as.integer(-1)
test_df[flag_cols][is.na(test_df[flag_cols])] <- as.integer(-1)

# Impute mean for one numeric column
train_df$White.Blood.cell.count..thousand.per.microliter.[is.na(train_df$White.Blood.cell.count..thousand.per.microliter.)] <-
  mean(train_df$White.Blood.cell.count..thousand.per.microliter., na.rm = TRUE)
test_df$White.Blood.cell.count..thousand.per.microliter.[is.na(test_df$White.Blood.cell.count..thousand.per.microliter.)] <-
  mean(test_df$White.Blood.cell.count..thousand.per.microliter., na.rm = TRUE)

```



```{r}
#lapply(train_df, unique)
```

### Pre-processing

###### Feature Pre-processing (including variable types/formats, names)

```{r feature_updates}
# preprocess variables
factor_cols <- c("Genes.in.mother.s.side",
                 "Inherited.from.father",
                 "Maternal.gene",
                 "Paternal.gene",
                 "Respiratory.Rate..breaths.min.",
                 "Heart.Rate..rates.min",
                 "Follow.up",
                 "Gender",
                 "Autopsy.shows.birth.defect..if.applicable.",
                 "Folic.acid.details..peri.conceptional.",
                 "H.O.serious.maternal.illness",
                 "H.O.radiation.exposure..x.ray.",
                 "H.O.substance.abuse",
                 "Assisted.conception.IVF.ART",
                 "History.of.anomalies.in.previous.pregnancies",
                 "Birth.defects",
                 "Blood.test.result",
                 "Disorder.Subclass",
                 "Genetic.Disorder")
train_df[factor_cols] <- lapply(train_df[factor_cols], factor)
test_df[factor_cols] <- lapply(test_df[factor_cols], factor)
# Note factors can be changed in dummy variables for better performances while in modeling.

# Generate updated summary of base dataset
str(train_df)
```

###### Collinearity and Dependencies

```{r collinearity_and_dependencies}
# Calculate Cramer's V "measure of association" between nominal factor variables (uses Chi-square statistic)
cscorr <- PairApply(train_df[ , sapply(train_df, is.factor)], CramerV, symmetric = TRUE)

# Shorten variable names for ease of reviewing output matrix
rn <- rownames(cscorr)
for (n in 1:length(rownames(cscorr))) {
  rn[n] <- paste(rownames(cscorr)[n], " (", AscToChar(64 + n), ")", sep = "")
  rownames(cscorr)[n] <- paste(AscToChar(64 + n))
}
for (n in 1:length(colnames(cscorr)))
  colnames(cscorr)[n] <- paste(AscToChar(64 + n))

# Show master list of variable names along with output ("correlation") matrix
cat(rn, sep = "\n")
cscorr
```

### Exploratory Data Analysis(EDA)

```{r}
dim(train_df)
```

###### Outlier detection

```{r Boxplots}
par(mar=c(10,2,1,1))
boxplot(train_df, las=2, col = c("turquoise","skyblue"))
```

**Note: Our data looks good on as there seems to be no outliers in the data.**

###### Frequency distribution

```{r Histogram}
pred_for_hist <- train_df[,2:29]
pred_for_hist <- pred_for_hist %>% mutate_if(is.character, as.numeric)
pred_for_hist <- pred_for_hist %>% mutate_if(is.factor, as.numeric)
par(mfrow = c(3, 3))

for (i in 1:ncol(pred_for_hist)) {
  hist(pred_for_hist[ ,i], xlab = names(pred_for_hist[i]), main = paste(names(pred_for_hist[i]), "Histogram"), col="blue")  
}
```

```{r density plot}
#par(mfrow = c(4, 4))
#for (i in 1:ncol(pred_for_hist)) {
#d <- density(pred_for_hist[,i], na.rm = TRUE)
#plot(d, main = paste(names(pred_for_hist[i]), "Density"))
#polygon(d, col="blue")
#}
```


###### Relation with target based on our hypothesis that the maternal and paternal genes might be cause of transmission of the genetic diorders. 

```{r}
library(ggplot2)
#install.packages("patchwork")
#install.packages("cowplot")

library(patchwork)
library(cowplot)
p1 <- ggplot(train_df, aes(x = train_df$Patient.Age, fill = train_df$Disorder.Subclass)) +geom_bar() + theme_classic()+scale_fill_hue(c=60, l=80)
p2 <- ggplot(train_df, aes(x = train_df$Maternal.gene, fill = train_df$Disorder.Subclass)) +geom_bar() +theme_classic()+ scale_fill_hue(c=60, l=80)
p3 <- ggplot(train_df, aes(x = train_df$Paternal.gene, fill = train_df$Disorder.Subclass)) +geom_bar() +theme_classic()+ scale_fill_hue(c=60, l=80)

p1+p2/p3
```


### Modeling

###### Assigning target and Predictors

```{r}

#test train before dummy 
train_x <- train_df[,2:29]
train_y <- train_df[,1]

test_x <- test_df[,2:29]
test_y <- test_df[,1]

dim(train_x)
dim(test_x)

```

```{r}
#subsetting numeric columns and character for changing categorical into dummy
library(dplyr)
# Subset numeric columns with dplyr

numeric_pred_train <- select_if(train_x, is.numeric)
numeric_pred_test<- select_if(test_x, is.numeric)

# Subset categorical columns with dplyr
cat_pred_train <- select_if(train_x,is.factor)
cat_pred_test <- select_if(test_x,is.factor)

dim(numeric_pred_train)
dim(numeric_pred_test)

dim(cat_pred_train)
dim(cat_pred_test)

```

```{r}
#encode to dummy
library(lattice)
dummies <- dummyVars(~ ., data=cat_pred_train[,1:17])
dummy_cat_df <- predict(dummies, cat_pred_train[,1:17])

dummies_test <- dummyVars(~ ., data=cat_pred_test[,1:17])
dummy_cat_df_test<- predict(dummies, cat_pred_test[,1:17])

#ready to model train test 
train_x<- as.data.frame(cbind(numeric_pred_train, dummy_cat_df))
train_y <- as.factor(train_y)

test_x<- as.data.frame(cbind(numeric_pred_test, dummy_cat_df_test))
test_y <- as.factor(test_y)

```

##### 

**Setting control function for our multiclass classification**

```{r}
ctrl <- trainControl(method = "cv",
                     summaryFunction = multiClassSummary,
                     classProbs = TRUE,
                     savePredictions = TRUE)
```

```{r nnet}

set.seed(476)

#nnetGrid <- expand.grid(size=1:3, decay=c(0,0.1,0.2,0.3,0.4,0.5,1,2))

#nnetFit <- train(x =  train_x, 
               # y = train_y,
              #  method = "nnet",
              #  tuneGrid = nnetGrid,
              #  metric = "ROC",
             #   trace = FALSE, 
              #  maxit = 2000, 
               # trControl = ctrl)

```



###### SVM (Support Vector machine )
```{r}

#sigmaEst <- kernlab::sigest(as.matrix(train_x))
#svmgrid <- expand.grid(sigma = sigmaEst, C = 2^seq(-4,+4))

#set.seed(476)
#svmRFit <- train(x = train_x , 
               # y = train_y,
               # method = "svmRadial",
               # tuneGrid = svmgrid,
               # metric = "ROC",
                #trControl = ctrl)
```






