---
title: "Predicting Genetic Disorders"
author:
  Emma Oo^[University of San Diego, eoo@sandiego.edu], Sindhu Bhattarai^[University of San Diego, sbhattarai@sandiego.edu], Dave Friesen^[University of San Diego, dfriesen@sandiego.edu]
date: "06/27/2022"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output:
  html_document:
    css: "style.css"
  pdf_document: default
---

<style>
.main-container {
  max-width: 1024px;
}
</style>


### Objective and Hypothesis


```{r setup, echo = FALSE, message = FALSE}
# Load R libraries

# Expand output width and minimize exp notation
options(width = 110)
options(scipen = 100)
options(digits = 4)

# Set style defaults
knitr::opts_chunk$set(fig.width = 5, fig.height = (5 * .7))
```


### Data Load and Validation

```{r data_load}
# Load datasets
gd_train_df <- read.csv("../data/train_genetic_disorders.csv", header = TRUE)
gd_test_df <- read.csv("../data/test_genetic_disorders.csv", header = TRUE)

# Data validation and understanding, including structure, content, and statistical characteristics covered below
```

#### Data (Sample) Characteristics (including variable types, formats, and names)

```{r data_chacteristics}
#
```

#### Categorical Variables

```{r categorical_variables}
#
```


### Data Splitting

```{r data_splitting}
#
```


### Data Cleaning (and reduction)

#### Zero/Near-Zero Variances

```{r near_zero_z_variances}
#
```

#### Missing Values

```{r missing_values}
#
```

#### Duplicate Values

```{r duplicate_values}
#
```

#### "Noisy" Data

```{r noisy_data}
#
```


### Data Transformation

#### Centering/Scaling (standardizing/normalizing)

```{r centering_scaling}
#
```

#### Statistical Characteristics (including distribution, skewness, outliers)

```{r statistical_characteristics}
#
```

#### Class Imbalances

```{r class_imbalances}
#
```

#### Other Feature Engineering (transformation, aggregation, enrichment)

```{r other_feature_engineering}
#
```


### Multivariate Analysis (and reduction)

#### Collinearity and Dependencies

```{r collinearity_and_dependencies}
#
```

#### Predictor Transformations (e.g., PCA)

```{r predictor_transformations}
#
```


### Modeling

#### Feature Selection

```{r feature_selection}
#
```

#### Training, Testing (validating), and Evaluation (iteration n)

```{r training_testing_evaluation}
#
```

#### Optimization, Tuning, Selection

```{r optimization_tuning_selection}
#
```
